\documentclass[english]{article}
	\usepackage[latin9]{inputenc}
	\usepackage{geometry}
	\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2cm}
	\setlength{\parindent}{0bp}
	\usepackage{mathrsfs}
	\usepackage{amsmath}
	\usepackage{setspace}
	\usepackage{esint}
	\usepackage{marginnote}
	\usepackage{hyperref}
	\usepackage{mathpazo} % font %
	\usepackage{eulervm} % font %
	\usepackage{MnSymbol}
	\usepackage{xcolor}

	\newcommand\cyan[1]{{\color{blue}{#1}}}
	\newcommand\orange[1]{{\color{orange}{#1}}}
	\newcommand\red[1]{{\color{red}{#1}}}
	\newcommand\task{\red{Task. }}
	\newcommand{\qed}{$\hfill\blacksquare$}
\begin{document}
\section*{0}
The author believes the following.
\\\\
A mathematician is an explorer, playfully discovering how things work, with his/her own methods and understanding of the mathematical world. As an extension of this, in this book, what one can be reasonably expected to discover themselves is not spelled out, only guided towards. \\\\
Except for brilliant insights which can appear to just come out of the blue, the reader must be motivated (i.e. feel intrigued) to explore an idea. As a corollary, proofs should often precede theorems (simply a reasoning with fruits).\\\\
An explorer should not be given as many tools as the world can offer - at least not initially. It is best to explore, perhaps naively, in order to develop an intuition and have a better appreciation of the more advanced tools to come later that make everything easier. As a corollary, concrete scenarios precede abstractions.

\newpage


\section*{1 - A first introduction to prime numbers}
\cyan{Some natural numbers, such as $6=2\cdot 3$, $9=3\cdot 3$, $15=3\cdot 5$, $24=2\cdot 12=3\cdot 8$, $51=3\cdot 17$ and $91=7\cdot 13$ may be decomposed as a product of smaller terms. The remaining numbers, including $1,2,3,5,7,11,13,17,19,\dots,93,97,\dots$ cannot be factored to smaller terms.}
\\\\
Definition. Let $n,k$ be integers, $0<k$. We say $k$ is a factor[divisor] of $n$ if $n=kr$ for some integer $r$. We write this relation as $k\mid n$, and also say $n$ is divisibly by $k$.
\\\\
\cyan{For given $k$, the integers divisibly by $k$ are $0,k,2k,3k,4k,5k,\dots$ as well as $-k,-2k,-3k,-4k,-5k,\dots$.}
\\\\
Understand. We have
\begin{itemize}
\item $1\mid k$ and $k\mid k$.
\item $k\mid r$ and $r\mid n$ imply $k\mid n$.
\item $k\mid n$ and $k\mid m$ imply $k\mid m\pm n$.
\item $k\mid r$ and $n\mid m$ imply $kn\mid rm$.
\item Given any pair of integers $n,k$ with $k>0$, there exists a unique representation $n=kt+s$ with $0\le s<k$.
\item $k\mid n$ implies that either $n=0$ or $k\le |n|$.
\\
\end{itemize} 
\cyan{Decomposing a natural number into a product of factors is analogous to decomposing physical matter into its elements. A water molecule is built of two hydrogen atoms and one oxygen atom, is analogous to $12$ being built of two $2$'s and one $3$, with $12=2\cdot 2\cdot 3$. }
\\\\
Definition. A natural number $p>1$ whose only divisors are one and itself is called a prime number.
\\\\
\cyan{Namely, a prime number is a number one cannot further divide. In formula, $p>1$ is prime if $a\mid p$ implies $a=1$ or $a=p$.}
\\\\
Understand. Each integer $n>1$ can be decomposed as a product of prime numbers (not necessarily different).
\\\\
\cyan{Note that we can factor $24$ in several ways - $2\cdot 12$, $3\cdot 8$, $4\cdot 6$. Continuing to divide all numbers until it is no longer possible, however, we always arrive at the same factorization, $24=2\cdot2\cdot2\cdot3$. To prove the uniqueness of the prime decomposition for all numbers, we first return to the chemistry analogy. When we form a product by mixing two substances in some way, and encounter an atom (a basic element) in the product, then this atom must have come from one of the two substances. In the number world, we have the analogous theorem.}
\\\\
Theorem. If the product $mn$ is divisible by a prime number $p$, then it must be that either $p\mid m$ or $p\mid n$.
\\\\
\cyan{The first three prime numbers are $2,3,5$. For $p=2$, the theorem simply states that a pair of odd numbers cannot have an even product. Indeed, any two odd numbers are of the $1+2k,1+2\ell$, with product $1+2(\ell+k+2\ell k)$. For $p=3$, a number not divisible by $3$ is of either form $1+3k$ or $2+3k$. So we have to verify that given two numbers where each is of either of these forms, their product is also in of of these forms. For $p=5$ there's already $4$ options for each form, so $16$ cases we have to check.}
\\\\
Understand. An equivalent formulation of above theorem is the following.
\\\\
\cyan{Theorem. Let $p$ be a prime number and $m,n$ be in $\{1,\dots,p-1\}$. Then $mn$ is not a product of $p$.}
\\\\
Recommendation. Try to find a proof of the theorem by yourself before reading the proof in the next page. Even if it takes a week. I believe in you. All the tools you need are in this page.
\newpage
\cyan{To motivate the proof, say for concreteness sakes $p=97$ and $m=15$. We have $p\mid 15n$, and we want to deduce that $97\mid n$. Write $15n=97k$. Clearly, $3\mid 97k$. We can safely assume (as we are now tackling the case $p=97$ that the theorem holds for the prime $3$, and so $3\mid 97$ or $3\mid k$. But $97$ is prime, so it must be that $3\mid k$. We can then divide both sides of the equation by $3$ to yield $5n=97j$, a similar equation, only $15$ was decreased to $5$. We can do one more step of this argument, now with the prime $5$, to reach the desired $97\mid n$.}
\\\\
Proof. Let $p$ be a given prime number, and \red{suppose the theorem holds for all smaller prime numbers.} Fix a number $n$ for which there exists a number $1\le m$ where $m$ is not divisible by $p$ but $mn$ is divisible by $p$, say $mn=pk$. Our goal is to show that $p\mid n$, namely to show that "$m$ can be replaced with $1$". For this, \red{we shall assume the $m$ we picked is the smallest of all its possible values}. Our goal is to now show that $m=1$. By the minimality of $m$, we have that $m$ is in the range $1,\dots,p-1$. If $m$ were bigger than $1$, then $m$ would have a prime divisor $q$. Moreover, $q$ would be smaller than $p$ (because $m$ is smaller than $p$), so that $q\mid mn=pk$ implies $q$ divides either $p$ or $k$. Of course, $q$ cannot divide the prime $p$, so $q$ divides $k$. Writing $m=qr$, $k=qd$, we have reached the equation $rn=pd$, but now $1\le r < m$ is a contradiction the minimality of $m$. This finishes the proof. \qed
\\\\
\cyan{Let us sketch a second proof. Again, suppose $p=97$ and $m=15$. Then $97\mid 15n$, and we want to deduce that $97\mid n$. We notice it is easier to increase $15$ as follows: we must have $97\mid 30n$, then $97\mid 90n$. But this yields $97\mid 7n$, and we replaced $15$ by the smaller $7$. We can repeat this argument to make the $7$ smaller, etc.}
\\\\
\task Make the above sketch into a rigorous proof.
\\\\
Understand. If $p$ is a prime number and $a_1,a_2,\dots,a_k$ is any sequence of integers with $p\mid a_1a_2\dots a_k$, then $p\mid a_j$ for some $1\le j\le k$.
\\\\
\cyan{The reason $2$ appears exactly $3$ times in the prime factorization $24=2\cdot2\cdot2\cdot3$ is because we can divide $24$ by $2$ three times, but no more. Namely, $24/2=12$, $12/2=6$, $6/2=3$, and $3$ cannot be divided by $2$. More generally.}
\\\\
Definition. Let $n$ be a positive integer and $p$ a prime number. "The number of times $n$ can be divided by $p$" is called the $p$ valuation of $n$ and denoted $\nu_p(n)$. Explicitly, $\nu_p(n)$ is the maximal non-negative integer $k$ for which $p^k\mid n$.
\\\\
Understand. As a corollary of the above theorem, if $n>1$ is an integer which decomposes as a product of (not necessarily distinct) primes $n=p_1p_2\dots p_k$, then for each prime number $p$, the number of times it appears in this factorization is precisely $\nu_p(n)$.


\newpage

\section*{? - lines and circles}
\cyan{In this chapter, we will see that often, lines and circles are one and the same. This might come as a surprise, and so my first goal to informally convince you that "lines are just circles with infinite radius". We shall give several ways to see this.}
\\\\
Clearly, taking a circle of radius $R$ around the origin and increasing $R$ does not get us close to anything but infinity. But imagine taking a circle passing through the origin, whose center is located at $(0,R)$. Taking $R$ to infinity and standing near the origin, it would appear the circles converge to the $x$ axis.
\href{https://www.desmos.com/calculator/wnpht6jrmp}{Here} is a desmos visualization.
\\\\
In your experience of riding a bicycle or driving a car in a circular path, you may have noticed this can be done by keeping the steering wheel at a constant spot, and not changing your speed. To trace a bigger circular path, the wheel needs to be closer to straight. Of course, a when the wheel is straight, you travel in a straight line.
\\\\
When a point travels at a constant pace around a circle, the direction of its motion is tangent to the circle. \red{If this is surprising to you, or you are not sure what this means, approximate the circle with a perfect $n$-gon inside it, with $n=10^{100}$, and think of the motion there instead.} In total, if the circle had radius $R$ meters and it took us $2\pi R$ seconds to make a full lap, then during those $2\pi R$ seconds the direction of motion changed at a constant pace of $2\pi$ radians per $2\pi R$ seconds, or at $1/R$ radians per second. Taking $R$ to infinity, $1/R$ goes to zero. A motion where the direction in which we are headed does not change is a motion along a straight line.
\\\\
Here is a final argument with physics (or calculus). Imagine again a particle traveling along a circle of radius $R$ meters, at a constant rate, completing a full lap after $2\pi R$ seconds. Then the particle's acceleration is towards the circle's center, at a magnitude of $1/R$ (meter per second squared). Taking $R$ to infinity, $1/R$ goes to zero. A motion with zero acceleration is a motion at a constant pace along a straight line.
\\\\
\cyan{We will now discuss some transformations that map circles and lines to each other. The first such map is known as a stereographic projection. Imagine a plane passing containing the equator of the earth. Now imagine that for each point $P$ on earth, except for the north pole $N$, we draw a line from $N$ to $P$. The sterographic projection of $P$ is the intersection of this line with the plane. \href{https://www.desmos.com/calculator/yrlyyh1kvn}{Here} is a desmos visualization in one dimension less.
%Notice that we can extend the stereographic projection continuosly to map the north pole to "the point at infinity". This makes sense, provided that there's one point at infinity, which is reached when traveling an infinite distance in any direction. Under this notion, it makes sense to think of every line (or plane) as containing the point at infinity.
}
\\\\
\task Let $\Phi$ be the stereogrphic projection of the sphere to the plane. Show that for every circle $C$ on the sphere passing through the north pole $N$, its image in the plane under $\Phi$ is a straight line. Moreover, this mapping is a bijection between circles on the sphere passing through the north pole and straight lines in the plane.

\end{document}
